---
title: "."
author: "Emma Rand"
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2: default
bibliography: ["refs/refs-r.bib", "refs/refs-project.bib"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.retina = 3)
```

```{r packages}
library(tidyverse)
library(patchwork)
library(likert)

```

```{r}
# palette
pal <- viridisLite::viridis(6)
```


# Introduction


# Methods

I used R [@R-core] with tidyverse packages [@tidyverse-pkg; @tidyverse-paper] for all analyses and the rmarkdown [@markdown-pkg; @markdown-book] and bookdown [@bookdown-pkg; @bookdown-book] packages for manuscript preparation.

```{r import-precourse-survey}
# data not shared
file <- "data-presurvey/responses.csv"
survey <- read_csv(file) 
```


```{r values}
participants <- survey$Name
n_partic <- length(participants)
```


```{r import-attend}
# data not shared
files <- dir("data-participation/", pattern = ".csv", full.names = TRUE)

attend <- files %>% 
  map_dfr(read_csv) %>% 
  filter(Name %in% participants) %>% 
  group_by(Name, session) %>% 
  summarise(minutes = sum(`Duration (Minutes)` ),
            attend = status) %>% 
  distinct()

```

# Results

```{r summarise-precourse}
analysis_summary <- survey %>% 
  group_by(`In what package/programme are you MOST likely to carry out data analysis and visualisation currently?`) %>% 
  summarise(n = length(`In what package/programme are you MOST likely to carry out data analysis and visualisation currently?`),
            mean_expertise = mean(`Please rate your expertise in R`),
            se_expertise = sd(`Please rate your expertise in R`)/sqrt(n)) 

report_summary <- survey %>% 
  group_by(`In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?`) %>% 
  summarise(n = length(`In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?`),
            mean_expertise = mean(`Please rate your expertise in R`),
            se_expertise = sd(`Please rate your expertise in R`)/sqrt(n))

```

```{r extract}
anal_r <- analysis_summary %>% 
  filter(`In what package/programme are you MOST likely to carry out data analysis and visualisation currently?` == "R") %>% 
  select(n) %>% as.numeric()

anal_not_r <- n_partic - anal_r
```

```{r}
attend %>% 
  ggplot() +
  geom_tile(aes(y = Name, x = session, fill = attend), colour = "white") +
  scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)]),
                    name = "Status")
```

There were `r n_partic` PhD students in the cohort.

```{r}

```



See Figure \@ref(fig:analysis-software-fig)

(ref:analysis-software-fig) Responses to the precourse survey question "In what package/programme are you MOST likely to carry out data analysis and visualisation currently?"


```{r analysis-software-fig, fig.cap='(ref:analysis-software-fig)'}
f1 <- analysis_summary %>% 
  ggplot(aes(x = reorder(`In what package/programme are you MOST likely to carry out data analysis and visualisation currently?`, n),
             y = n)) +
  geom_bar(fill = "black", stat = "identity") +
  scale_x_discrete(name = "",
                   expand = c(0, 0)) +
  scale_y_continuous(name = "Number of people",
                     expand = c(0, 0),
                     limits = c(0, 16),
                     breaks = seq(0, 15, 5),) +
  theme_classic() + 
  theme(plot.margin = margin(t = 0, r = 0, l = 0)) +
  coord_flip()

f2 <- analysis_summary %>% 
  ggplot(aes(x = reorder(`In what package/programme are you MOST likely to carry out data analysis and visualisation currently?`,n), 
             y = mean_expertise)) +
  geom_bar(fill = "white", stat = "identity") +
  geom_point() +
  geom_errorbar(aes(x = `In what package/programme are you MOST likely to carry out data analysis and visualisation currently?`,
                    ymin = mean_expertise - se_expertise,
                    ymax = mean_expertise + se_expertise),
                width = 0.3) +
  scale_x_discrete(name = "",
                   labels = NULL,
                   expand = c(0, 0)) +
  scale_y_continuous(name = "Self-rated expertise with R",
                     expand = c(0, 0),
                     breaks = seq(0, 10, 2),
                     limits = c(0, 10.5)) +
  theme_classic() + 
  theme(plot.margin = margin(t = 0, r = 0, l = 0)) +
  coord_flip()

layout <- "
ab
"

f1 + f2 + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))
  
```

See Figure \@ref(fig:report-software-fig)

(ref:report-software-fig) Responses to the precourse survey question "In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?"


```{r report-software-fig, fig.cap='(ref:report-software-fig)'}
f1 <- report_summary %>% 
  ggplot(aes(x = reorder(`In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?`,n),
             y = n)) +
  geom_bar(fill = "black", stat = "identity") +
  scale_x_discrete(name = "",
                   expand = c(0, 0)) +
  scale_y_continuous(name = "Number of people",
                     expand = c(0, 0),
                     limits = c(0, 46),
                     breaks = seq(0, 40, 10),) +
  theme_classic() + 
  theme(plot.margin = margin(t = 0, r = 0, l = 0)) +
  coord_flip()

f2 <- report_summary %>% 
  ggplot(aes(x = reorder(`In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?`, n), 
             y = mean_expertise)) +
  geom_bar(fill = "white", stat = "identity") +
  geom_point() +
  geom_errorbar(aes(x = `In what package/programme are you MOST likely to write up analysis results to submit to a journal or similar?`,
                    ymin = mean_expertise - se_expertise,
                    ymax = mean_expertise + se_expertise),
                width = 0.3) +
  scale_x_discrete(name = "",
                   labels = NULL,
                   expand = c(0, 0)) +
  scale_y_continuous(name = "Self-rated expertise with R",
                     expand = c(0, 0),
                     breaks = seq(0, 10, 2),
                     limits = c(0, 10.5)) +
  theme_classic() + 
  theme(plot.margin = margin(t = 0, r = 0, l = 0)) +
  coord_flip()

layout <- "
ab
"

f1 + f2 + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))
 
```
```{r}
pre_comfort <- survey %>% 
  select(starts_with("Please rate your level of comfort" )) %>%
  drop_na() %>% 
  data.frame(check.names = FALSE)


comfort_q <- "Please rate your level of comfort with the following in R"

names(pre_comfort) <- names(pre_comfort) %>% 
  sub(pattern = paste0(comfort_q, "\\s+\\["), 
      replacement = "") %>% 
  sub(pattern = "\\]", replacement = "")         

levels <- c("Not at all comfortable",
            "A bit uncomfortable",
            "Quite comfortable",
            "Completely comfortable")

pre_comfort <- pre_comfort %>% 
  mutate(`working directories, relative and absolute paths` =
           factor(`working directories, relative and absolute paths`, 
                  levels = levels) ,
         `importing data in plain text files` =
           factor(`importing data in plain text files`, 
                  levels = levels) ,
         `importing data in almost any file format` =
           factor(`importing data in almost any file format`, 
                  levels = levels) ,
         `importing data through APIs` =
           factor(`importing data through APIs`, 
                  levels = levels) ,
         `reshaping data` =
           factor(`reshaping data`, 
                  levels = levels),
         `summarising data` =
           factor(`summarising data`, 
                  levels = levels),
         `text processing column names and cell contents` =
           factor(`text processing column names and cell contents`, 
                  levels = levels),
         `plotting data` =
           factor(`plotting data`, 
                  levels = levels),
         `formatting figures for publication` =
           factor(`reshaping data`, 
                  levels = levels),
         `the tidyverse` =
           factor(`the tidyverse`, 
                  levels = levels),
         `the pipe %>%` =
           factor(`the pipe %>%`, 
                  levels = levels),
         `RMarkdown for blending text, analysis and figures` =
           factor(`RMarkdown for blending text, analysis and figures`, 
                  levels = levels),
         `More advanced RMarkdown features such as cross referenced tables, figures, citations and a wide variety of YAML options ` =
           factor(`More advanced RMarkdown features such as cross referenced tables, figures, citations and a wide variety of YAML options `, 
                  levels = levels))
```

```{r}
comfort <- likert(pre_comfort)
order <- names(pre_comfort)

plot(comfort,
          group.order = order,
          legend.position = "top",
          legend = "",
          text.size = 3,
          low.color = pal[3],
          high.col = pal[5]) +
  theme(axis.title.x = element_blank(),
        axis.line.x = element_blank())
```

See Table \@ref(tab:summary-table).  

```{r summary-table}

```



# Discussion


# References

